---
title: "Visualize SFS"
author: "Bruno de Medeiros"
date: "2/29/2020"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths('~/vdiR/library/')
library(tidyverse)
```

In this script, we will visualize the site frequency spectra that we will use to infer migration rates using fastsimcoal. 

Let's start by reading the SFS files for Anchylorhynchus 1, the species on both S. coronata and S. botryophora.

```{r}
species = 'Anchylorhynchus_1'

dimensions = scan(paste0(species,'/',species,'_MSFS.obs'),
                  what='integer',
                  sep=' ',
                  skip = 1,
                  nlines = 1) %>%
  as.numeric

dimensions = dimensions[-1]


folded_counts = scan(paste0(species,'/',species,'_MSFS.obs'),
                  what='integer',
                  sep=' ',
                  skip = 2,
                  nlines = 1) %>%
  as.numeric 


if (length(dimensions) == 3){
  df = expand.grid(freq_3 = 0:dimensions[3],
                 freq_2 = 0:dimensions[2],
                 freq_1 = 0:dimensions[1]) %>%
  mutate(count = folded_counts) 
} else {
  df = expand.grid(freq_2 = 0:dimensions[2],
                 freq_1 = 0:dimensions[1]) %>%
  mutate(count = folded_counts)
}


df

```

Now we can use group_by to marginalize. Let's look at Population 1 vs population 2.

```{r}
df_temp = df %>%
  group_by(freq_1,freq_2) %>%
  summarize(count = sum(count)) %>%
  mutate(new_freq_1 = ifelse(freq_1+freq_2>(sum(dimensions[1:2])/2),
                             max(freq_1)-freq_1,
                             freq_1),
         new_freq_2 = ifelse(freq_1+freq_2>(sum(dimensions[1:2])/2),
                             max(freq_2)-freq_2,
                             freq_2)
         ) %>%
  group_by(new_freq_1, new_freq_2) %>%
  summarise(count = sum(count)) %>%
  rename(freq_1 = new_freq_1,
         freq_2 = new_freq_2)



ggplot(df_temp) +
  geom_raster(aes(x = freq_1, y = freq_2, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)
```

Now pop 1 vs pop 3:

```{r}
df_temp = df %>%
  group_by(freq_1,freq_3) %>%
  summarize(count = sum(count)) %>%
  mutate(new_freq_1 = ifelse(freq_1+freq_3>(sum(dimensions[c(1,3)])/2),
                             max(freq_1)-freq_1,
                             freq_1),
         new_freq_3 = ifelse(freq_1+freq_3>(sum(dimensions[c(1,3)])/2),
                             max(freq_3)-freq_3,
                             freq_3)
         ) %>%
  group_by(new_freq_1, new_freq_3) %>%
  summarise(count = sum(count)) %>%
  rename(freq_1 = new_freq_1,
         freq_3 = new_freq_3)



ggplot(df_temp) +
  geom_raster(aes(x = freq_1, y = freq_3, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)
```

Now pop 2 vs pop 3:
```{r}
df_temp = df %>%
  group_by(freq_2,freq_3) %>%
  mutate(new_freq_2 = ifelse(freq_2+freq_3>(sum(dimensions[c(2,3)])/2),
                             max(freq_2)-freq_2,
                             freq_2),
         new_freq_3 = ifelse(freq_2+freq_3>(sum(dimensions[c(2,3)])/2),
                             max(freq_3)-freq_3,
                             freq_3)
         ) %>%
  group_by(new_freq_2, new_freq_3) %>%
  summarise(count = sum(count)) %>%
  rename(freq_2 = new_freq_2,
         freq_3 = new_freq_3)

ggplot(df_temp) +
  geom_raster(aes(x = freq_2, y = freq_3, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)
```

Now let's transform this in a function to visualize all species. 

```{r}
show_SFS = function(species){
  
  dimensions = scan(paste0(species,'/',species,'_MSFS.obs'),
                  what='integer',
                  sep=' ',
                  skip = 1,
                  nlines = 1) %>%
  as.numeric

dimensions = dimensions[-1]


folded_counts = scan(paste0(species,'/',species,'_MSFS.obs'),
                  what='integer',
                  sep=' ',
                  skip = 2,
                  nlines = 1) %>%
  as.numeric 


if (length(dimensions) == 3){
  df = expand.grid(freq_3 = 0:dimensions[3],
                 freq_2 = 0:dimensions[2],
                 freq_1 = 0:dimensions[1]) %>%
  mutate(count = folded_counts) 
  
p1 = df %>%
  mutate(new_freq_1 = ifelse(freq_1+freq_2>(sum(dimensions[1:2])/2),
                             max(freq_1)-freq_1,
                             freq_1),
         new_freq_2 = ifelse(freq_1+freq_2>(sum(dimensions[1:2])/2),
                             max(freq_2)-freq_2,
                             freq_2)
         ) %>%
  group_by(new_freq_1, new_freq_2) %>%
  summarise(count = sum(count)) %>%
  rename(freq_1 = new_freq_1,
         freq_2 = new_freq_2) %>%
  ggplot() +
  geom_raster(aes(x = freq_1, y = freq_2, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)

p2 = df %>%
  mutate(new_freq_1 = ifelse(freq_1+freq_3>(sum(dimensions[c(1,3)])/2),
                             max(freq_1)-freq_1,
                             freq_1),
         new_freq_3 = ifelse(freq_1+freq_3>(sum(dimensions[c(1,3)])/2),
                             max(freq_3)-freq_3,
                             freq_3)
         ) %>%
  group_by(new_freq_1, new_freq_3) %>%
  summarise(count = sum(count)) %>%
  rename(freq_1 = new_freq_1,
         freq_3 = new_freq_3) %>%
  ggplot() +
  geom_raster(aes(x = freq_1, y = freq_3, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)

p3 = df %>%
  mutate(new_freq_2 = ifelse(freq_2+freq_3>(sum(dimensions[c(2,3)])/2),
                             max(freq_2)-freq_2,
                             freq_2),
         new_freq_3 = ifelse(freq_2+freq_3>(sum(dimensions[c(2,3)])/2),
                             max(freq_3)-freq_3,
                             freq_3)
         ) %>%
  group_by(new_freq_2, new_freq_3) %>%
  summarise(count = sum(count)) %>%
  rename(freq_2 = new_freq_2,
         freq_3 = new_freq_3) %>%
  ggplot() +
  geom_raster(aes(x = freq_2, y = freq_3, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)

print(list(p1,p2,p3))
  
} else {
  df = expand.grid(freq_2 = 0:dimensions[2],
                 freq_1 = 0:dimensions[1]) %>%
  mutate(count = folded_counts)
  
  ggplot(df) +
  geom_raster(aes(x = freq_1, y = freq_2, fill = count)) +
  scale_fill_viridis_c(option = 'B', trans='log1p', breaks = 10^(1:6)) +
  coord_cartesian(expand=F)
}
  
}
```

Let's apply the function to each species:
```{r}
show_SFS('Anchylorhynchus_1')
```

```{r}
show_SFS('Anchylorhynchus_2')
```

```{r}
show_SFS('Andranthobius_1')
```

```{r}
show_SFS('Andranthobius_2')
```

```{r}
show_SFS('C_decolor_1')
```

```{r}
show_SFS('C_decolor_2')
```

```{r}
show_SFS('C_impar')
```


```{r}
show_SFS('D_polyphaga')
```

```{r}
show_SFS('M_bondari_1')
```

```{r}
show_SFS('M_ypsilon')
```

```{r}
show_SFS('R_rectinasus_1')
```